---
title: "GLM"
author: "Richard J. Telford"
output: html_document
---

```{r, messages = 'hide'} 
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises GLM

## Poisson GLM
1. Import sorich.csv
```{r data import}
# Importing data
sorich.df <- read.csv("sorich.csv")
```
<br><br> 
2. Fit a linear model for species richness (nsp) against cover assuming a Normal distribution
```{r fit linear model}
# Fitting model (adapted from lecture)
fit_lm <- lm(nsp ~ cover, data = sorich.df)

# Checking p-values for model - looks good, model is statistically significant
summary(fit_lm)
```
<br><br> 
3. Study the diagnostic plots
```{r Diagnostic plots}
# Diagnostic plots 
par(mfrow = c(2,2))
plot(fit_lm)
par(mfrow = c(1,1))
# Normal Q-Q plot does not look so good
```
<br><br> 
*Normal Q-Q plot does not look so good, not sure about scale-location.*
<br><br> 
4. What distribution is the response?
```{r plot of nsp ~ cover}
# Plotting nsp ~ cover to see distribution 
library(tidyverse)
ggplot(sorich.df, aes(x = cover, y = nsp)) + 
  geom_point() + 
  labs(y = "Species Richness", x = "Cover") +
  theme_bw(base_size = 16) 
```
<br><br> 
*The distribution of the data is not normal (gaussian), but is an example of a poisson-distributed dataset.* 
<br><br> 
5. Do an appropriate analysis
```{r glm}
# Doing an appropriate analysis
# trying with glm and poisson distribution because data is cover aka proportion
fit_glm_poi <- glm(nsp ~ cover, data = sorich.df, family = poisson)

# Checking p-values for model 
summary(fit_glm_poi) # looks good, model is statistically significant
```
<br><br> 
6. Check for over-dispersion. If necessary modify the model.
```{r overdispersion}
# Running anova to check for overdispersion 
anova(fit_glm_poi)
# Residual df < Residual deviance - overdispersed 
#         690 < 1315.0 --> res. dev. 2x more than res. df 
```
<br>  
7. Interpret the results <br> 
*Resudial deviance is 2x more than residual df, i.e. the model is overdispersed*
<br><br>  
8. How does the width of the confidence interval at cover = 10 change when over-dispersion is allowed for
```{r plot, confidence interval ++}
# TRIED PLOTTING ON MY OWN, DID NOT TURN OUT CORRECT
# Plot of cover ~ species richness
ggplot(sorich.df, aes(x = cover, y = nsp)) +
  geom_point() +
  labs(y = "Species Richness", x = "Cover") +
  theme_bw(base_size = 16) +
# Adding model to coverplot to see if it fits tha data
  geom_abline(data = fit_glm_poi, color = "hotpink") # STRIAGHT LINE NOT CURVED :(

# PLOT + LINE FROM LECTURE: 
ggplot(cbind(sorich.df,
             fit = fitted(fit_glm_poi)), # make sure you fit the correct model
       aes(x = cover, y = nsp)) + 
  geom_point() + 
  geom_line(aes(y = fit), colour = "royalblue") +
  labs(y = "Species Richness", x = "Cover", title = "Q8: Plot with model line")

```

<br><br>  
9. Do the grasses differ from the herb? <br> 
i.e. make grassherb a factor and include it in the model and test its significance.
<br><br> 
```{r grasses and herbs}
# Checking names in sorich.df 
names(sorich.df) # grasherb already a name? 

# Making grasherb a factor(object?) to test its significance in the model 
grasherb <- sorich.df["grasherb"] 
str(grasherb) 

# Including it in my model 'fit_glm_poi'
fit_glm_gh <- glm(nsp ~ cover * grasherb, data = sorich.df, family = poisson)
summary(fit_glm_gh) # it seems like grasherb is a significant term/interaction/factor
anova(fit_glm_gh)   # still overdispersed (i think?)
```

<br><br> 

## How much does over-dispersion affect results
1. Use `sample()` to randomise the response in sorich to remove the relationship between the predictor and response.
2. Test if a Poisson GLM with cover as a predictor is significant
3. Repeat 1000 times and find the distribution of p-values

## Binomial GLM and GAM

1. Open library mgcv
2. Import data pot.csv
3. What type of distribution is the response variable?
4. What type of link-function do we use?
5. Do an appropriate GLM analysis?
6. Interpret the results?
7. Do a GAM analysis
8. Compare the GLM and GAM models.
9. Which model would you prefer, and why?