---
title: "Practical 3"
author: "Richard J. Telford"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## GLS

With the built-in `iris` data.
```{r Data import}
# Loading libraries
library(tidyverse)

# Importing data 
iris.df <- iris # built in dataset, no need to import 
str(iris.df)
```


Make a plot to show how `Petal.Length` varies between species.
```{r Plot of petal length and species}
# Plotting petal length to see if it varies between species 
ggplot(data = iris.df, aes(x = Species, y = Petal.Length, fill = Species)) +
  geom_boxplot() + 
  theme_bw(base_size = 20) +
  labs(title = "Boxplot: Petal length and species", x = "Species", y = "Petal Length")
```

Find the variance of `Petal.Length` for each species.
```{r Variance of petal length}
# Finding variance for each species 
iris.df %>% 
  group_by(Species) %>% 
  summarise(
  variance = var(Petal.Length)) # 'variance' is the name of 'var(Petal.Length)'
```


Fit an anova using `lm` between `Petal.Length` and species and examine the diagnostic plots.
```{r Fitting anova, diagnostic plots}
# Making linear model to investigate xx
fit.lm_petal.sp <- lm(Petal.Length ~ Species, data = iris.df)
plot(fit.lm_petal.sp) # 

# Not sure if this is necessary, but tried it 
anova(fit.lm_petal.sp)
summary(fit.lm_petal.sp) 
```

Fit a `gls` for the same model. Have the coefficients changed? (remember to load ??nlme?? package)
```{r Generalized least squared for my model `fit.lm_petal.sp`}
# EXPLAIN WHY
library(nlme)
fit_gls1_petal.sp <- gls(Petal.Length ~ Species, data = iris.df)
summary(fit_gls1_petal.sp)
```


Fit a `gls` for the same model but allow the variance to be different for each species by adding `varIdent` to the `weights` argument of the model. 
```{r Variance different for each species}
# 
fit_gls2_petal.sp <- gls(Petal.Length ~ Species, data = iris.df, 
                        weights = varIdent(form = ~ +1|Species))
summary(fit_gls2_petal.sp)
```

Use `AIC` to test if this is a better model.
