---
title: "Linear models"
author: "Richard J. Telford"
date: "May 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear Modelling Exercises

### 1. Import the penguin data from the `palmerpenguin` package
```{r Importing packages and datasets}
# Importing libraries needed
library(palmerpenguins)
library(tidyverse)

# Importing data from 'palmerpenguins' package
penguins.df <- palmerpenguins::penguins # pulls out data from the palmerpenguins 
                                        # package and creates object 'penguins.df'
```

### 2. Do big birds have big beaks? Fit a linear model between bill length and body mass, and interpret the coefficients.
```{r Linear model and plot}
# Fitting model to see if big birds have big beaks
fit.lm <- lm(bill_length_mm ~ body_mass_g, data = penguins.df)
summary(fit.lm)

# Making a plot to see the data
ggplot(penguins.df, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point() +
  theme_bw(base_size = 20) +
  labs(x = "Body mass (g)", y = "Bill length (mm)", 
       title = "Do big birds have big beaks?") +
  stat_smooth(method = "lm", col = "hotpink") # did not work 

```

**Interpretation of coefficients** <br>
*As the t values are large this looks very good. A large t value tell you that there are few/little errors in the data. Multiple R-squared is 0.35, and thus this factor/predictor explains 35% of the variation of the trait.* 

### 3. Examine the diagnostics plots. Is everything OK?
```{r Diagnostic plots}
# Examining the diagnostic plots of my model - fit.lm
par(mfrow = c(2,2))
plot(fit.lm)
par(mfrow = c(1,1))
```

**Interpretation of diagnotsic plots** <br>
*The diagnostic plots look ok* <br>
 * Residuals vs fitted: relatively flat <br> 
 * Normal Q-Q: ok, not great at the 'upper' end, not on the line <br> 
 * Scale-location: ok, not great, should be flatter <br>
 * Residuals vs Leverage: looks great <br>

### 3. Calculate the residual sum of squares (RSS) (find sum the squared difference between the estimates calculated from the coefficients and the observed values). <br> 
```{r Calculate the residual sum of squares (RSS)}
# Finding the coefficients in my model (instead of copy-pasting from summary)
# to calculate residual sum of squares (RSS)

# Coefficients in my model 'fit.lm'
coef(fit.lm)

# Coefficient 1 (Intercept) from my model 'fit.lm'
coef(fit.lm)[1] # (Intercept) 26.89887

# Coefficient 2 (body_mass_g) from my model 'fit.lm'
coef(fit.lm)[2] # body_mass_g 0.004051417

# Residual sum of squares (RSS)
# This is the best model, and we can see that by changing the numbers in them 
# and see that they are not smaller i.e. not a better 'calculation' that the RSS
sum((penguins.df$bill_length_mm - ((coef(fit.lm)[1]) + ((coef(fit.lm)[2]) * penguins.df$body_mass_g)))^2, na.rm = TRUE)
# 6564.494
```
<br>
<br>
### 4. Recalculate the Residual sum of squares using different values for the coefficients (in the formula yi = Beta + Beta xi + ??i). <br>
```{r Recalculate the Residual sum of squares}
# Changing 'coef(fit.lm)[1]' to the number 3 to recalculate the RSS
sum((penguins.df$bill_length_mm - ((3) + ((coef(fit.lm)[2]) * penguins.df$body_mass_g)))^2, na.rm = TRUE)
# 201899.9 > 6564.494
```

<br> 
<br>
5. Is the new RSS larger or smaller?<br>
**The new RSS is (much) bigger than the previous one from the linear model.** <br>
6. Estimate the slope using covariance and variance.
```{r Covariance and variance}
# Finding stuff i need to calculate cov and var

# Variance of bill length and body mass
var_bl <- var(penguins.df$bill_length_mm, na.rm = TRUE) 
var_bm <- var(penguins.df$body_mass_g, na.rm = TRUE) 

# Covariance of bill length and body mass
covar_bl_bm <- cov(x = penguins.df$body_mass_g, 
                   y = penguins.df$bill_length_mm,
                   use = 'pairwise.complete.obs')
```


7. Do you get the same result?
8. Add species to the model as a predictor.
9. Does species improve the model?
10. How should the coefficients be interpreted?

