---
title: "Bio302 Practical 1 Exploratory Data Analysis"
author: "Richard J. Telford"
date: "May 25, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1 Peguins

Load the `palmerpenguins` package and see the `penguins` data.
```{r}
#install.package("palmerpenguins")
library(palmerpenguins)
library(tidyverse)
penguins.df <- palmerpenguins::penguins
str(penguins.df) # view did not work
names(penguins.df) # checking out the names, they look ok 

penguins.df <- na.omit(palmerpenguins::penguins)
```


Find the range, mean, median and variance of variable `bill_length_mm`.
```{r}
# Calculating range, mean, median and variance of variable `bill_length_mm` 
# through making new separate objects
bill_length.range    <- range(penguins.df$bill_length_mm, na.rm = TRUE)
bill_length_mean     <- mean(penguins.df$bill_length_mm, na.rm = TRUE)
bill_length_median   <- median(penguins.df$bill_length_mm, na.rm = TRUE)
bill_length_variance <- var(penguins.df$bill_length_mm, na.rm = TRUE)

# Trying to use the summarise function to calculate 
# range, mean, median and variance of variable `bill_length_mm`
penguins.df %>% summarise(
  mean = mean(bill_length_mm), # 'mean' is the name of 'mean(bill_length_mm)'
  median = median(bill_length_mm), # 'median' is the name of 'median(bill_length_mm)'
  range = range(bill_length_mm), # 'range' is the name of 'range(bill_length_mm)'
  variance = var(bill_length_mm)) # 'variance' is the name of 'var(bill_length_mm)'

```


What is the standard error of the mean of `bill_length_mm`.
```{r}
# Before calculating the standard error i will calculate the standard deviation
# Calculating standard deviation of the mean of `bill_length_mm`
sd.bill_length_mean <- sqrt(bill_length_variance)

# Calculating standard error of the mean of `bill_length_mm`
se.bill_length_mean <- sd.bill_length_mean/sqrt(bill_length_mean)
```

Plot a histogram of variable `bill_length_mm`.
```{r}
# Simple histogram of bill length (mm)
ggplot(penguins.df, aes(x = bill_length_mm)) +
  geom_histogram(fill = 'gray90') + # color of fill of bars
  theme_bw() + # theme black and white (i think)
  labs(title = "Bill length - Palmer's Penguins", # title of plot
              x = "Bill length (mm)", y = "Count")# labels of x- and y-axis
```

Mark the position of the mean and median.
```{r}
# Simple histogram of bill length (mm) with mean and median 
ggplot(penguins.df, aes(x = bill_length_mm)) +
  geom_histogram(fill = 'gray90') + # color of fill of bars
  theme_bw(base_size = 20) + # theme black and white (i think)
  labs(title = "Bill length - Palmer's Penguins", # title of plot
              x = "Bill length (mm)", y = "Count") + # labels of x- and y-axis
  # adding the mean in form of a blue line
  geom_vline(xintercept = bill_length_mean, linetype = 1, color = 'royalblue', size = 1) +
  # # adding the median in form of a  line
  geom_vline(xintercept = bill_length_median, linetype = 1, color = 'hotpink', size = 1) 
```

Plot `bill_length_mm` against `bill_depth_mm`
```{r}
# Plot of bill length ~ bill depth 
ggplot(penguins.df, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(shape = 19, color = 'royalblue') + # color of fill of points
  theme_bw(base_size = 20) + # theme black and white (i think)
  labs(title = "Bill length against bill depth - Palmer's Penguins", # title of plot
              x = "Bill length (mm)", y = "Bill depth (mm)")  # labels of x- and y-axis
```

Find the covariance and correlation of `bill_length_mm` against `bill_depth_mm`.
```{r}
# Covariance of bill length and bill depth
covar_bill_length_depth <- cov(x = penguins.df$bill_length_mm, 
                               y = penguins.df$bill_depth_mm,
                               use = 'pairwise.complete.obs')

# Correlation of bill length and bill depth
corr_bill_length_depth <- cor(x = penguins.df$bill_length_mm, 
                               y = penguins.df$bill_depth_mm,
                               use = 'pairwise.complete.obs')
```

Is the correlation significant? (hint use cor.test())
```{r}
# xxx
cor.test(x = penguins.df$bill_length_mm, 
         y = penguins.df$bill_depth_mm,
         use = 'pairwise.complete.obs')
```
*Yes, the correlation is significant (p-value < 0.01).*

How should the p-value be interpreted?
*---*

There are three levels in factor `species` Plot `bill_length_mm` against `bill_depth_mm`, coloured by `species`
```{r}
# Plot of bill length ~ bill depth, coloured by species 
ggplot(penguins.df, aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_point(shape = 19) +
               #c('Adelie' = 'royalblue',
  #               'Gentoo' = 'seagreen3',
  #               'Chinstrap' = 'palevioletred')) + # color of fill of points
  theme_bw(base_size = 20) + # theme black and white (i think)
  labs(title = "Bill length ~ bill depth - Palmer's Penguins", # title of plot
              x = "Bill length (mm)", y = "Bill depth (mm)")  # labels of x- and y-axis
```

# <span style="color:red">**CONTINUE FROM HERE**</span>

Is the correlation between `bill_length_mm` against `bill_depth_mm` significant for any of the groups?
```{r}
# Making three objects with only the species to compare them to each other
species_1 <- filter(penguins.df, species == 'Adelie')
species_2 <- filter(penguins.df, species == 'Gentoo')
species_2 <- filter(penguins.df, species == 'Chinstrap')

# Comparing the groups to each other 

```

This is a sub-group analysis, what are the dangers of this type of analysis?

# Part 2 not penguins

Install the `datasauRus` package with `install.packages("datasauRus")` and load it with `library`.

Dataset `datasaurus_dozen` has `x` and `y` variables and a dataset name. 
For each dataset, calculate the mean and standard deviation of x and y and the correlationn between x and y.