---
title: "Power tests"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Power test app

Run `biostats.apps::power_lm_app()` and explore the effect of the number of observations on the width of the confidence intervals, the proportion of trials that are statistically significant, and the effect size of significant trials.

```{r}
# Running app to explore statistical stuff
biostats.apps::power_lm_app()
```

<br>
# Simulating a power test

A scientist is setting up an experiment to test whether a new type of food gives more weight gain in fish. They need to know how many fish they need to sample to have an 80% probability (i.e. power = 0.8) of getting a statistically significant result.

The experiment has a control (standard food) and a treatment arm (new food). <br>

```{r power test}
# General power test stuff from the lecture 'power tests': 
# Two-sample t test power calculation 
library("pwr")
pow <- pwr.t.test(d = 0.5, # effect size
                  sig.level = 0.05, 
                  power = 0.8) 
pow
plot(pow)
# need 64 observations in each group to reach statistical power of 0.8 
```


Fish typically have a mass of **1kg** (`delta = 1`) with a standard deviation of **0.2kg** (`sd = 2`).The scientist hopes that the new food will give at least 0.1kg extra mass (otherwise it is not worth the extra cost). <br>

**Simulate data for a single experiment with 10 fish in each arm of the experiment.** <br>
<span style="color: red;">**Test if your experiment was significant.**</span>
```{r simulation of fish experiment}
# From lectureslide 11/16 - WITH SLIGHT ALTERATIONS
# data info
n <- 10    # number observations in each group
delta <- 1 # difference between means 
sd <- 2    # standard deviation 

# simulate means
mu <- rep(c(0, delta), each = n)

# add noise
y <- mu + rnorm(length(mu), sd = sd)

## predictor
x <- factor(rep(c("A", "B"), each = n))

# Run test
test <- t.test(y ~ x) # not significant, too small sample size 
```


**Make your simulation into a function and repeat it 100 times.** <br> 
<span style="color: red;">**What proportion of trials are statistically significant?**</span> <br>
<span style="color: blue;">Have not found out :(</span>
```{r making function to rerun}
# MAKING A FUNCTION to rerun 
# from lecture notes: 
sim_t_test <- function(n, delta, sd, ...){
  # simulate means
  mu <- rep(c(0, delta), each = n)
  # add noise
  y <- mu + rnorm(length(mu), sd = sd)
  ## predictor
  x <- factor(rep(c("A", "B"), each = n))
  # run test
  test <- t.test(y ~ x)
  broom::tidy(test) %>% mutate(n = n, delta = delta, sd = sd) 
}

```

```{r running the simulation function we just made}
library("broom")
library("tidyverse")
# Lecture simulation - 30 fish 
sim_t_test(n = 10, # 30 fish 
           delta = 1, # minimal effect size we are looking for
           sd = 2) 

# repeat many times
nrep = 100

# crossing two matrices, not sure why 
control <- crossing(rep_no = 1:nrep, 
                    n = seq(10, 100, 20)) # from 10, to 100, series of 20

runs <- control %>%
  pmap_df(sim_t_test, # the simulation function we just made
          delta = 1, # minimal effect size we are looking for 
          sd = 2) %>% 
  mutate(sig = p.value <= 0.05) 

plot_30fish <- runs  %>%
  group_by(n) %>%
  summarise(power = mean(sig)) %>%
  ggplot(aes(x = n, y = power)) +
  geom_line() + 
  geom_point() 

# shows which p-values that are <0.05 (true) and which are >0.05
runs %>% 
  group_by(sig) %>% 
  summarise(n = n()) %>% 
  mutate(proportion = n*100/500)
# power = 63.4 - a bit too low 
```


Run the simulation for different numbers of fish. How many are needed to get enough power?
```{r my simulation(s)}
# DO NOT CHANGE THE NAME 'sim_t_test' - IT IS THE FUNCTION'S NAME
# My simulation - 100 fish 
sim_t_test(n = 2000, # 100 fish
           delta = 1, sd = 2)

# How many times (n) you want to repeat (rep) the simulation 
# (can take a while if you want to run it MANY times)
nrep = 120 

# ?
control <- crossing(rep_no = 1:nrep, # not quite sure whats going on here
                    n = seq(10, 120, 20)) # seq(from n, to n, series of n)

# Running the simulation 
runs <- control %>%
  pmap_df(sim_t_test, # the simulation function we just made
          delta = 1, # minimal effect size we are looking for 
          sd = 2) %>% # st dev 
  mutate(sig = p.value <= 0.05) # setting 

# shows which p-values that are <0.05 (true) and which are >0.05
runs %>% 
  group_by(sig) %>% 
  summarise(n = n()) %>% 
  mutate(proportion = n*100/720)
```


Plot your results.
```{r plotting results}
plot_100fish <- runs  %>% # plotting the 100-fish simulation 
  group_by(n) %>% # grouping by number of fish (100)
  summarise(power = mean(sig)) %>% # 
  ggplot(aes(x = n, y = power)) + #
  geom_line() + # line
  geom_point() # scatterplot 

plot_100fish # creating plot 
```


# A second variable

The scientist thinks that a second variable, temperature, is also important, so expands the experiment to include this second variable with two levels, cool and warm.
<br>
Extend your power analysis to cope with the new experimental design.
```{r 2nd var - extending power analysis}
# EXTENDING POWER ANALYSIS BY ADDING PREDICTOR VARIABLE h
# From lectureslide 11/16
# data info
n <- 100    # number observations in each group, changed it from 30 to 64
delta <- 0.1 # difference between means 
sd <- 0.2    # standard deviation 

# simulate means
mu <- rep(c(0, delta), each = n)
#?rep

# add noise
y <- mu + rnorm(length(mu), sd = sd)

# Making my predictor variables: x and h
# x for food type: A and B 
# each in function rep: A A A A A B B B B B B
x <- factor(rep(c("A", "B"), each = n)) 

# h for temperature: cool and warm
# times in function rep: cool warm cool warm cool warm 
h <- factor(rep(c("cool", "warm"), times = n)) 

# Run test - two-way-anova aka linear model because 2x categorical predictors 
temptest.h <- lm(y ~ h)         # temperature predictor 
temptest1 <- lm(y ~ x + h)      # additive model 

anova(temptest.h, temptest1)    # temptest1 lowest RSS, and statistically significant

# Checking 
temptest2 <- lm(y ~ x * h)      # interaction model 

anova(temptest1, temptest2) 
# temptest2 has slightly lower RSS, but it is not statiistically significant,
# so we keep/consider both temptest1, temptest2
```
<br>
How many fish do they need to get enough power?
```{r new power analysis}
# 



```


<br>
Plot your result

<br>
# Distribution of p values under Null hypothesis

Simulate some data where the null hypothesis is true (ie effect size is zero).

Plot the distribution of p-values. 

Compare this to the distribution of p-values calculated above.


