---
title: "GLM"
author: "Richard J. Telford"
output: html_document
---

```{r, message = 'hide', warning = FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

# Exercises GLM

## Poisson GLM
### 1. Import sorich.csv
```{r prep}
# Loading packages
library(tidyverse)
library(nlme)
library(mgcv)
# Importing data
sorich.df <- read.csv("sorich.csv")
```
<br><br> 
### 2. Fit a linear model for species richness (nsp) against cover assuming a Normal distribution
```{r fit linear model}
# Fitting model (adapted from lecture)
fit_lm <- lm(nsp ~ cover, data = sorich.df)

# Checking p-values for model - looks good, model is statistically significant
summary(fit_lm)
```
<br><br> 

### 3. Study the diagnostic plots
```{r Diagnostic plots}
# Diagnostic plots 
par(mfrow = c(2,2))
plot(fit_lm) # Normal Q-Q plot does not look so good
```

```{r settings, include = FALSE}
par(mfrow = c(1,1))
```

<br><br> 
*Normal Q-Q plot does not look so good, not sure about scale-location.*
<br><br> 

### 4. What distribution is the response?
```{r plot of nsp ~ cover}
# Plotting nsp ~ cover to see distribution 
library(tidyverse)
ggplot(sorich.df, aes(x = cover, y = nsp)) + 
  geom_point() + 
  labs(title = "Plot 1 - Cover and Species Richness \nPoisson distribution", 
       y = "Species Richness", x = "Cover") + 
  theme_bw(base_size = 16) 
```
<br><br> 
*The data is not normally distributed, but is an example of a poisson-distributed data.* <br>
*When the data is in proportions (percent covered) we normally perform(?) logistic regressian and use the function `glm()`* 
<br><br> 

### 5. Do an appropriate analysis.
```{r glm}
# Doing an appropriate analysis
# trying with glm and poisson distribution because data is cover aka proportion
fit_glm_poi <- glm(nsp ~ cover, data = sorich.df, family = poisson)

# Checking p-values for model 
summary(fit_glm_poi) # looks good, model is statistically significant
```
<br><br> 

### 6. Check for over-dispersion. If necessary modify the model.
```{r overdispersion}
# Running anova to check for overdispersion 
anova(fit_glm_poi)
# Residual df < Residual deviance - overdispersed 
#         690 < 1315.0 --> res. dev. 2x more than res. df 

# SHould it be like this instead?
anova(fit_glm_poi, test = "Chisq")
```
<br>  

### 7. Interpret the results <br> 
*Resudial deviance is 2x more than residual df, i.e. the model is overdispersed*
<br><br>  

### 8. How does the width of the confidence interval at cover = 10 change when over-dispersion is allowed for?
```{r plot, confidence interval - my code}
# TRIED PLOTTING ON MY OWN, DID NOT TURN OUT CORRECT
# Plot of cover ~ species richness
# ggplot(sorich.df, fit = fitted(fit_glm_poi), # borrowed some code from lecture slide
#        aes(x = cover, y = nsp)) +
#   geom_point() +
#   geom_line(aes(y = fit), colour = "hotpink")
#   labs(title = "Plot 2 - Cover and Species Richness 
#        \nPoisson distribution with model line", 
#        y = "Species Richness", x = "Cover") +
#   theme_bw(base_size = 16) +
# # Adding model to coverplot to see if it fits the data 
# #  geom_abline(fit_glm_poi, color = "hotpink") # STRIAGHT LINE NOT CURVED :(
```


```{r lecture plot and line, message = FALSE, warning = FALSE}
# PLOT + LINE FROM LECTURE: 
ggplot(cbind(sorich.df,
             fit = fitted(fit_glm_poi)), # make sure you fit the correct model
       aes(x = cover, y = nsp)) + 
  geom_point() + 
  # tried to add confidence intervals in 'geom_line' but it did not work
  geom_line(aes(y = fit), colour = "red" , fill = "grey", se = TRUE) + 
  # tried to add confidence intervals with code underneath, 
  # and managed to do so, but lines did not match
  geom_smooth(method = glm, colour = "royalblue", fill = "gray", se = TRUE) +
  labs(title = "Plot 3 - Cover and Species Richness \nPoisson distribution with model line(s)", 
       y = "Species Richness", x = "Cover") + 
  theme_bw(base_size = 16) 
```

<br><br>  
9. Do the grasses differ from the herb? <br> 
i.e. make grassherb a factor and include it in the model and test its significance.
<br><br> 
```{r grasses and herbs}
# Checking names in sorich.df 
names(sorich.df) # grasherb already a name? 

# Making grasherb a factor(object?) to test its significance in the model 
grasherb <- sorich.df["grasherb"] 
str(grasherb) 

# Including it in my model 'fit_glm_poi'
fit_glm_gh <- glm(nsp ~ cover * grasherb, data = sorich.df, family = poisson)
summary(fit_glm_gh) # it seems like grasherb is a significant term/interaction/factor
anova(fit_glm_gh)   # still overdispersed (i think?)
```

<br><br> 

## How much does over-dispersion affect results
**1. Use `sample()` to randomise the response in sorich to remove the relationship between the predictor and response.**
```{r randomise response, results='hide'}
# Not quite sure what Im doing or if it is right 
sample(sorich.df, size = 100, replace = TRUE)
```
<br><br>
**2. Test if a Poisson GLM with cover as a predictor is significant.**
*Help*
<br><br>
**3. Repeat 1000 times and find the distribution of p-values.**
*help again*
<br><br>
## Binomial GLM and GAM

**1. Open library mgcv**
**2. Import data pot.csv**
```{r importing data}
# Importing data
pot.df <- read_csv("pot.csv")
```

<br><br>
**3. What type of distribution is the response variable?**
```{r plot of pot}
# Plotting data so check its distribution
ggplot(pot.df, aes(x = alt, y = potalp)) +
  geom_point() +
  theme_bw(base_size = 16) +
  labs(title = "Binomial distribution")
```

<br><br>
**4. What type of link-function do we use?**
<br>
*We use a* ***logistic link function*** *because the data is* ***binomially distributed.***
<br><br>
**5. Do an appropriate GLM analysis.**
```{r GLM analysis}
#
```

<br><br>
**6. Interpret the results.**
<br><br>
**7. Do a GAM analysis.**
<br><br>
**8. Compare the GLM and GAM models.**
<br><br>
**9. Which model would you prefer, and why?**