---
title: "Spatial and temporal structure"
author: "Richard J. Telford"
date: "June 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = FALSE, message = FALSE)
```

1) Simulate a 100-observation autocorrelated timeseries with `arima.sim`, with a first order autoregressive coefficient of 0.5. Also make a time vector of `1:100`
```{r Simulation}
# Making simulation 
arima.sim(n = 100, )

ts.sim <- arima.sim(list
                    # specifying that this is an AR1-model, 1st order model
                    (order = c(1,0,0), 
                      # order = c(2,0,0) --> AR2-model, 2nd order model
                         ar = 0.5), # autoregressive coefficient
                    n = 100) # number of obs

time.vector <- 1:100
```


2) Plot the data. 
```{r Plot 1}
# Plotting the data
ts.plot(ts.sim)
```



3) Regress the timeseries against time with an OLS model. Does the model appear to be statistically significant?
```{r}
# Normally you would use lm() 
gls.1 <- gls(ts.sim ~ time.vector)
anova(gls.1)

# linear model 
lm1 <- lm(ts.sim ~ time.vector)
anova(lm1)
```


4) Plot the model diagnostics, including an `acf` and `pacf` of the residuals.
```{r Plot 2 - model diagnostic}
# 
plot(lm1)

# Autocorrelation function
acf(resid(lm1))

# Partial autocorrelation function
pacf(resid(lm1))
```


5) Use the Durbin-Watson test to test the residuals for autocorrelation.
```{r Durbin-Watson test}
# install.packages("lmtest")
library(lmtest)

# Durbin-Watson test to check for autocorrelation 
dwtest(lm1) 
# Values between 0-4, DW = 2 no autocorrelation 
#                         <2 positive autocorrelation 
#                         >2 negative autocorrelation 

# DW = 0.7918, p-value = 2.696e-10
# Positive autocorrelation, very significant p-value
# null hypothesis rejected i.e. there is positive autocorrelation in model lm1 
```
*DW = 0.7918, p-value = 2.696e-10* <br>
*Positive autocorrelation, very significant p-value* <br>
*Null hypothesis rejected i.e.* ***there is positive autocorrelation in model lm1*** <br>


6) Fit a gls with an appropriate correlation structure. Is this a better model? How have the p-value and effect size changed?
```{r}

```


7) Repeat the above 1000 times and find how autocorrelation affects the distribution of the p-value and the effect size.


## Real data

1) The built-in dataset LakeHuron has annual lake level data from 1885 to 1972
Load the data with the command data(LakeHuron)

2) Plot the data.

3) Regress the LakeHuron lake level against year using a linear model. Is there a significant trend?

4) Plot the autocorrelation and partial autocorrelation functions for the residuals from this regression. Interpret them.

5) Fit an autoregressive models to the residuals. Compare the results with your interpretation of the PACF plot.


6) Fit a gls model using a corAR1 correlation structure. Test if the correlation structure is necessary. Is the trend significant? 

7) Fit a gls model using a corARMA correlation structure with two AR terms. Is this model an improvement?


